---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---
```{r pkg-knitr-opts, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, fig.retina=2, message=FALSE, warning=FALSE)
options(width=120)
```

[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/iptrie.svg?branch=master)](https://travis-ci.org/hrbrmstr/iptrie) 
[![Coverage Status](https://codecov.io/gh/hrbrmstr/iptrie/branch/master/graph/badge.svg)](https://codecov.io/gh/hrbrmstr/iptrie)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/iptrie)](https://cran.r-project.org/package=iptrie)

# iptrie

Efficiently Store and Query 'IPv4' Internet Addresses with Associated Data

## Description

Tries are great ways to store data that has obvious hierarchichal 
properties, such as 'IPv4' addresses. Methods are provided to create 'IPv4'
tries and store, retrieve and delete 'IPv4' address keys with values. 
Functions are based on the 'zmap' 'iptree' 'C' library.

## NOTE

This is an experiment but will likely turn into a CRAN package or get migrated
to either the `iptools` or `asntools` package. 

I initially wanted to  feel the pain (again) of using R's own C interface
(i.e. no `Rcpp` crutch) since it's easy to forget just how handy Rcpp is. 
(It turns out that my ill memories of playing at the R C level are also unjustified).

Having said that, I also need this functionality and the basic functions here 
will have more robust conterparts soon.

PRs and issues welcome.

## What's Inside The Tin

The following functions are implemented:

- `as_iptrie`: Convert a data frame to an IP trie
- `iptrie`:  Efficiently Store and Query 'IPv4' Internet Addresses with Associated Data
- `iptrie_create`: Create a new IPv4 Trie
- `iptrie_destroy`:  Destroy an IP trie
- `iptrie_insert`: Insert a value for an IPv4 Address+CIDR combo into an IPv4 Trie
- `iptrie_ip_in`:  Lookup a value for an IPv4 Address+CIDR combo into an IPv4 Trie or Test for Existence
- `iptrie_lookup`: Lookup a value for an IPv4 Address+CIDR combo into an IPv4 Trie or Test for Existence
- `iptrie_remove`: Remove a trie entry for an IPv4 Address+CIDR combo into an IPv4 Trie
- `is_iptrie`: Create a new IPv4 Trie
## Installation

```{r install-ex, eval=FALSE}
devtools::install_git("https://sr.ht.com/~hrbrmstr/iptrie.git")
# or
devtools::install_git("https://gitlab.com/hrbrmstr/iptrie.git")
# or (if you must)
devtools::install_github("hrbrmstr/iptree")
```

## Usage

```{r lib-ex}
library(iptrie)
library(tidyverse)

# current version
packageVersion("iptree")

```

### Basic Usage

```{r basic}
x <- iptrie_create()

iptrie_insert(x, "10.1.10.0/24", "HOME")

iptrie_ip_in(x,"10.1.10.1/32")

iptrie_ip_in(x,"10.1.11.1/32")

iptrie_lookup(x, "10.1.10.1/32")

iptrie_lookup(x, "10.1.10.1/32", "exact")
```

### Data frame to iptrie

```{r dftotrie}
xdf <- data.frame(a = "10.1.10.0/24", b = "HOME", stringsAsFactors = FALSE)

(xt <- as_iptrie(xdf))

is_iptrie(xt)

(xt <- as_iptrie(xdf, "a", "b"))

iptrie_ip_in(xt, "10.1.10.6")
```

### Bigger example (kinda what `as.data.frame.iptrie` does)

```{r bigger, cache=TRUE}
# Make a trie from autonomous system CIDRs
xdf <- astools::routeviews_latest()

cat(scales::comma(nrow(xdf)), "\n")

asntrie <- iptrie_create()

system.time(for (i in 1:nrow(xdf)) {
  iptrie_insert(asntrie, xdf[["cidr"]][i], xdf[["asn"]][i])
})

# Get a block list (picked at ransom)
blklst <- url("https://iplists.firehol.org/files/botscout_1d.ipset")
y <- readLines(blklst)
close(blklst)

y <- y[!grepl("^#", y)] # comments at the top

cat(scales::comma(length(y)), "\n")

system.time(do.call(
  rbind.data.frame,
  lapply(y, function(.x) {
    r <- iptrie_lookup(asntrie, .x, "best")
    if (is.null(r)) {
      data.frame(
        ip = .x, 
        asn = NA_character_, 
        cidr = NA_character_,
        stringsAsFactors = FALSE
      )
    } else {
      data.frame(
        ip = .x, 
        asn = r, 
        cidr = sprintf("%s/%d", attr(r, "ip"), attr(r, "mask")),
        stringsAsFactors = FALSE
      )
    }
  })
) -> cdf)

as_tibble(cdf)

count(cdf, cidr, sort=TRUE)
```

## iptree Metrics

```{r cloc, echo=FALSE}
cloc::cloc_pkg_md()
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
